---
title: "Exploratory Data Analysis"
author: Alison Elgass
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(ggridges)
```

# Load in & clean/tidy data
(cache = true means it won't re-download on each knit)
```{r cache=TRUE}
#THIS WILL TAKE A WHILE
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728",
                               "USC00519397",
                               "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>% 
  
  mutate(name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
       tmin = tmin / 10,
       tmax = tmax / 10, 
  month = lubridate::floor_date(date, unit = "month")) %>%
  
  select(name, id, month, everything())
```


# Using 'group_by' and 'count'
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(), #number of entries
    n_unique = n_distinct(date))
```

Use 'count' instead to do the same thing as above
```{r}
weather_df %>% 
    count(month) #like saying group by name, then count

weather_df %>% 
    count(month, name = "n_days") #specify name of new col
```

Let's make a nice table
```{r}
weather_df %>% 
  count(name) %>% 
  knitr::kable() #MAKES IT PRETTY!!!
```

## 2 x 2 tables, a digression...

Two ways to do it (but Jeff is anti 2x2 table)
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(                    #create new column 'cold'
    cold = case_when(        #using this new fxn
      tmax < 5  ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ "")
    ) %>% 
  
  group_by(name, cold) %>% 
  count() %>% 
  pivot_wider(
    names_from = cold,
    values_from = n
  )
  
```

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate( 
    cold = case_when(
      tmax < 5  ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ "")
    ) %>% 
  
  janitor::tabyl(name, cold)
```

# Some summaries
```{r}
weather_df %>% 
  group_by(name, month) %>% #group by 3 locations
  summarize( #create cols: count, mean, sd, median
    n = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE),
    median_pprcp = median(prcp, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
           geom_point() + geom_line()
```

"Un-tidy" using pivot_wider to make it more reader-friendly
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax)
    ) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
    ) %>% 
  knitr::kable(digits = 1)
```

## Grouped mutates

(don't forget you can ungroup...)
```{r results = FALSE}
weather_df %>% 
  group_by(name) %>% 
  ungroup()
```

```{r}
#same df (still 1095 rows), but new column 'mean_tmax' is by group (so all Central Park rows have same mean)
weather_df %>% 
  group_by(name, month) %>% 
  mutate(                             #create new columns
    mean_tmax = mean(tmax, na.rm = TRUE), #group-specific
    centered_tmax = tmax - mean_tmax      
  ) %>% 
  view
```

window functions in grouped mutates
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    tmax_rank = min_rank(tmax) #rank by coldest max temp
                               #in each location
  ) %>%                        #in each month
  filter(tmax_rank == 1) %>%   #now show #1's (3)
  view
```

logs and leads...
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax), #yesterday's tmax
    one_day_change = tmax - lagged_tmax
  ) %>% 
  summarize(sd_daily_change = 
              sd(one_day_change, na.rm = TRUE)) %>% 
  view
```


