---
title: "Visualization 2"
author: Alison Elgass
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

# Load in & clean/tidy data
Cache=true means it won't re-download on each knit
```{r cache=TRUE}
#THIS WILL TAKE A WHILE
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728",
                               "USC00519397",
                               "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>% 
  
  mutate(name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
       tmin = tmin / 10,
       tmax = tmax / 10) %>% 
  
  select(name, id, everything())
```

# Add labels & ticks
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) +
 scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15")
  ) +
  scale_color_hue(
    name = "Weather Station",
    h = c(50, 200)
 )
```


# Create base plot, use Viridis Package
```{r}
#This time we "save" the plot as a variable to use later
base_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

base_temp_plot #now display it
```

## Themes (ORDER MATTERS)
```{r}
base_temp_plot +
  theme_minimal() + #many theme options!
  theme(legend.position = "bottom")
  
```

### Side note on warnings
Warning message: remove 15 rows containing missing values
It's ok, ggplot leaves them out of plot
```{r}
#see these 15 rows have missing values 
weather_df %>% 
  filter(is.na(tmin))
```

# Using patchwork package for multiple panels
```{r}
#Filter data by location, create 3 separate df's
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, 
       aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_line(data = central_park)
```

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```


# Data Manipulation
Using forcats & factors (more examples online)
```{r}
weather_df %>%
  mutate(
    name = forcats::fct_relevel(
      name, 
      c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))
  ) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color="blue", alpha=0.5) + 
  theme(legend.position = "bottom")
```

## Restructure then plot
```{r}
weather_df %>%
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp"
  ) #NOT WORKING!!!!!!!!!!!!!!!

  ggplot(weather_df, aes(x = temp, fill = observation)) +
    geom_density(alpha = .5)

```



